\large Fowizzle: Combining spectra or images by forward modeling

### 2024-03-09 paper/main.tex
this isn't interpolation; flux conservation?

% To-Do items
% -----------
% - What's our name? Choose and propagate into where "Forward Model(tm)" is used.
% - Put in notes about our hypothesis for WHY we get uncorrelated pixels.
% - Make comments directly about Drizzle to justify the name.
% - discuss conservation of flux vs intensity.
% - Blanton asks: Is this identical to sinc shift and average? Hogg thinks it MIGHT be.
% - Add Hogg response to Blanton: Interpolate with (say) cubic spline and then average is NOT THE SAME as fit all the data simultaneously with a cubic spline function on a particular set of knots.
% - Dustin asks: Is this effectively producing a sinc LSF? Hogg thinks it DEFINITELY IS NOT. Explain why not in the text.
% - Hogg notes: Is it clear in the text that we are taking everything to a standard instrument? And the user can decide what that instrument is? That isn't in our current implementation, but it COULD BE. Discussion point?
% - Related Hogg issue: Any latent spectrum (like our combined spectrum) should be something that COULD BE OBSERVED by some instrument in the future. Hence spectro-perfectionism is wrong (ish).
% - Read and cite the DESI spectral pipeline paper; Dustin says it's relevant.
% - Audit for rigid terminology. HOGG is currently liking ``combined spectrum'' for the model expectation, for example. And "standard practice" for the traditions of our field.
% - Audit and fix the sign applied to $\Delta x_i$ everywhere. Wrong relative to code?
% - Write about the two input-data cases.
% - See bug list in code for other issues with the text--code relationship (noise model; line-formation model).
% - Audit for its / it's. Hogg has issues with this.
% - Deal with all HOGG, ANDY, ARC, XXX, YYY, CITE, etc.

@@

It would be wrong to say that the practice of astronomy is, basically, \emph{staring at the sky}.
That said, staring at the sky is a big part of what we do.
And the way we do it, usually, is this:
We take many individual images of our target with exposure times that are long enough to make read noise irrelevant (if possible) while at the same time short enough to avoid saturation (of what we care about).
We then shift and average (maybe median?) those images to get the highest possible signal-to-noise on our target.
This, fundamentally, is the technology underlying the \textsl{Hubble Deep Field} (\citealt{hdf}),
the \textsl{APOGEE} spectroscopic survey (\citealt{apogee}),
and the future Rubin Observatory \textsl{LSST} (\citealt{lsst}), among countless other projects.
This technology is encoded in tools such as \textsl{Drizzle} (\citealt{drizzle}) and HOGG.

@@

@@

We have presented a forward-modeling approach to combining spectra or images.
This method has good properties:
It doesn't involve interpolating or otherwise distorting the input data;
it produces a combined or mean spectrum in the format that traditional methods employ;
it makes no assumptions about the form of the combined spectrum (except that it can be represented on an $M_\ast$-pixel grid);
and it has excellent noise properties in real circumstances.
The method can work on a wide variety of different kinds of data, but it is especially valuable in projects where the individual input exposures are ill-sampled, or not as well sampled as one might like.
It is also good for situations in which the input data have heterogeneous bad pixels or data gaps.

>>>>

% To-Do items
% -----------
% - What's our name? Choose and propagate into where "Forward Model(tm)" is used.
% - Put in notes about our hypothesis for WHY we get uncorrelated pixels.
% - Make comments directly about Drizzle to justify the name.
% - discuss conservation of flux vs intensity.
% - Hogg notes: Is it clear in the text that we are taking everything to a standard instrument? And the user can decide what that instrument is? That isn't in our current implementation, but it COULD BE. Discussion point?
% - Related Hogg issue: Any latent spectrum (like our combined spectrum) should be something that COULD BE OBSERVED by some instrument in the future. Hence spectro-perfectionism is wrong (ish).
% - Read and cite the DESI spectral pipeline paper; Dustin says it's relevant.
% - Audit for rigid terminology. HOGG is currently liking ``combined spectrum'' for the model expectation, for example. And "standard practice" for the traditions of our field.
% - Audit and fix the sign applied to $\Delta x_i$ everywhere. Wrong relative to code?
% - Write about the two input-data cases.
% - See bug list in code for other issues with the text--code relationship (noise model; line-formation model).
% - Deal with all HOGG, ANDY, ARC, XXX, YYY, CITE, etc.

@@

It would be wrong to say that the practice of astronomy is, basically, \emph{staring at the sky}.
That said, staring at the sky is a big part of what we do.
And the way we do it, usually, is this:
We take many individual images of our target with exposure times that are long enough to make read noise irrelevant (if possible) while at the same time short enough to avoid saturation (of what we care about).
We then shift and average (maybe median?) those images to get the highest possible signal-to-noise on our target.
This, fundamentally, is the technology underlying the \textsl{Hubble Deep Field} (\citealt{hdf}),
the \textsl{APOGEE} spectroscopic survey (\citealt{apogee}),
and the future Rubin Observatory \textsl{LSST} (\citealt{lsst}), among countless other projects.
This technology is encoded in tools such as \textsl{Drizzle} (\citealt{drizzle}).

@@

\paragraph{Flux or photon conservation}
HOGG SHOULD WE PUT WORDS HERE ABOUT WHY WE DON'T CARE ABOUT THIS CONSIDERATION? (1) normalized spectra. (2) It is intensity that a telescope imager measures.

@@

We have presented a forward-modeling approach to combining spectra or images.
This method has good properties:
It doesn't involve interpolating or otherwise distorting the input data;
it produces a combined or mean spectrum in the format that astronomers traditionally expect;
it makes no assumptions about the form of the combined spectrum (except that it can be represented on an $M_\ast$-pixel grid);
and it has excellent noise properties in real circumstances.
The method can work on a wide variety of different kinds of data, but it is especially valuable in projects where the individual input exposures are ill-sampled, or not as well sampled as one might like; our original motivation was the individual \textsl{APOGEE} spectrograph (\citealt{apogeehardware}) exposures.
It is also good for situations in which the input data have heterogeneous bad pixels or data gaps.

The method does not involve interpolating the data.
However, it does fit a very flexible model to the data, and there is a limit in which fitting very flexible models is equivalent to interpolation (see, eg, \citealt{fittingflexible}).
So is this method, implicitly, interpolating the data?
Our answer is no.
Interpolation of the data involves non-parametric models with control points that are set by the pixel grids of the input data.
Our model choice never refers to the properties of the input data pixel grids; it is chosen based on the desired properties of the output, combined-spectrum grid.
Furthermore, interpolation functions (like cubic spline, say, or Lanczos-5, say) pass through every raw-data pixel value exactly.
There is no point in this method at which the function being fit passes through every (or even any) of the raw-data pixel values.
